Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2024-03-13 2_SSL_NE/article/article_supplementary.tex
updated figures

\clearpage

\section{Previous Literature on Techological Progress in LEDs}
\label{sec:prev_lit}

@@

\section{Metrics: Technological Progress in LEDs}

Investigating the sources of rapid progress in LED technology over the past decades, in which white LEDs have come to dominate the lighting market \cite{zissis2021}, requires selecting appropriate metrics for tracking and quantifying this progress. The choice of metrics affects both what data sources can be used in the analysis and which research methodologies can be used to calculate and analyse such metrics. We select the metrics based on the following two general criteria. First, we focus only on metrics widely accepted and reported in industry because metrics proposed in the scientific literature but not reported by device manufacturers cannot be used to compare the performance of commercial LED devices over time. Second, the chosen metrics must be useful for understanding the impact of individual technological improvements on relevant performance and cost characteristics.

Historically, the progress in LED technology has been commonly described by pointing to impressive improvements in LED device performance, more specifically device brightness, as well as electrical efficiency, and manufacturing cost reductions \cite{Taki2019}. However, metrics related to consumer experience from using white LEDs as a lighting source, such as the perceived temperature of a white light source and its ability to faithfully render colours, have also played a significant part in the market adoption of new lighting technologies \cite{Menanteau2000,Sandahl2006,CAIRD2008,murphy2012governing} and received substantial attention in LED research and development efforts \cite{azevedo2009transition,cho2017white}. Therefore, a comprehensive analysis of the evolution in white LED technology must take into consideration advances in 1) physical device performance; 2) consumer experience; and 3) LED device manufacturing costs. Next, we introduce and discuss the metrics that we use to track progress in each of these three areas.

\section{Metrics: Device Performance}
\label{sec:device_performance_metrics}

@@

\clearpage
\section{Metrics: Consumer Experience}

>>>>

\clearpage
\section{Previous Literature on Technological Progress in LEDs}
\label{sec:prev_lit}

@@

\clearpage
\section{Metrics for Technological Progress in LEDs}

\textbf{Motivation for Methods Selection}

Investigating the sources of rapid progress in LED technology over the past decades, in which white LEDs have come to dominate the lighting market \cite{zissis2021}, requires selecting appropriate metrics for tracking and quantifying this progress. The choice of metrics affects both what data sources can be used in the analysis and which research methodologies can be used to calculate and analyse such metrics. We select the metrics based on the following two general criteria. First, we focus only on metrics widely accepted and reported in industry. This is because metrics proposed in the scientific literature, but not reported by device manufacturers, cannot be used to compare the performance of commercial LED devices over time. Second, the chosen metrics must be useful for understanding the impact of individual technological improvements on relevant performance and cost characteristics.

Historically, the progress in LED technology has been commonly described by pointing to impressive improvements in LED device performance, more specifically device brightness, as well as electrical efficiency, and manufacturing cost reductions \cite{Taki2019}. However, metrics related to the quality of light, such as the perceived temperature of a white light source and its ability to faithfully render colours, have also played a significant part in the market adoption of new lighting technologies \cite{Menanteau2000,Sandahl2006,CAIRD2008,murphy2012governing} and received substantial attention in LED research and development efforts \cite{azevedo2009transition,cho2017white}. Therefore, a comprehensive analysis of the evolution in white LED technology must take into consideration advances in 1) physical device performance; 2) consumer experience; and 3) LED device manufacturing costs. Next, we introduce and discuss the metrics that we use to track progress in each of these three areas.

\textbf{Distinction between Metrics}

We use a well-established distinction between metrics. Those metrics which are related to the spectrum of the light source only ("color quality metrics" \cite{DLC_SSL_Requirements}\cite{dilaura2011lighting}), we group under the term consumer experience metrics. Those metrics, which are determined by the number and energy of photons leaving the device (as a function of the number of electrons entering it), we group under the term device performance metrics. As such, our distinction is physical at the fundamental level. This follows existing literature on the technological progress in solid-state lighting \cite{tsao2010solid}. We describe these metrics in detail in \cref{sec:device_performance_metrics}-\cref{sec:metrics_consumer}.

\textbf{"Haitz's Law"}

A highly cited and frequently called-upon metric is \textit{flux/package}, as described in what has become known as Haitz's law. An observation rather than a law of nature, it describes increasing device performance and decreasing manufacturing cost. Originally introduced in 2000 \cite{haitz1999case}, it has been updated frequently since then. Unfortunately, the way Haitz's law is frequently used is  problematic due to the misleading conclusions it suggests. While the original version presented by Haitz displays the increasing performance of light-emitting diode chips, the graphs have been extended with datapoints from lamps which contain an unspecified number of chips. The metric therefore loses its value in describing technological progress at the chip-level. The continuing rise in performance inferred from this chart is disconnected from the actual, much lower, performance of single chips \cite{weinold2021compound}. This is shown in \cref{fig:haitz-org}-\cref{fig:haitz}.

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{./figures/haitz_law_old_both.pdf}
    \caption{Historical increase in flux for white light-emitting diode packages (\textit{"Haitz's Law"}), as it is often reproduced in literature. The original data by Haitz for red LEDs from 2000 is indicated for reference. Note the different range on the abscissa compared to \cref{fig:haitz}. Sources for red data: \cite{haitz1999case}\cite{haitz2011solid}. Sources for white-light data: figure \cref{fig:haitz}.}
    \label{fig:haitz-org}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{./figures/haitz_law_white.pdf}
    \caption{Historical increase in flux for white light-emitting diode chips and packages, inspired by \textit{"Haitz's Law"}. Note that the increase in flux per single chip is not increasing as starkly as the flux for multi-chip packages. Shown are datapoints for best commercial performers from press releases, datasheets and and industry periodicals. \\ Sources for cost: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid} \\ Sources for flux: \cite{haitz2007seoul}\cite{haitz2010osram}\cite{haitz2010lumi-1}\cite{haitz2010cree} \cite{haitz2011lumi-1}\cite{haitz2014nichia}\cite{haitz2015cree-1}\cite{haitz2015cree-2}\cite{haitz2018cree}\cite{haitz2019osram}\cite{haitz2019lumi-1}\cite{haitz2019lumi-1}\cite{haitz2020lumi-1} \cite{haitz2013cob}\cite{haitz2016cob}\cite{haitz2017cob}\cite{haitz2019cob}\cite{haitz2019lumi-1}\cite{haitz2019lumi-2}}
    \label{fig:haitz}
    \vspace{-20mm}
\end{figure}

\clearpage
\section{Metrics: Device Performance}
\label{sec:device_performance_metrics}

@@

\clearpage
\section{Metrics: Consumer Experience}
\label{sec:metrics_consumer}
